---
title: "ECO1400 Final Paper Reg"
Date: "Dec 2022"
output: pdf_document
urlcolor: blue
geometry: margin=0.75in
---

Install and load required R packages
```{r, message=FALSE}
if (!require("Ecdat")) install.packages("Ecdat")
if (!require("LINselect")) install.packages("LINselect")
if (!require("gmm")) install.packages("gmm") # function gmm
if (!require("vars")) install.packages("vars")
if (!require("xtable")) install.packages("xtable")
if (!require("StepReg")) install.packages("StepReg")
if (!require("corrplot")) install.packages("corrplot")
if (!require("tidycensus")) install.packages("tidycensus")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("plyr")) install.packages("plyr")
if (!require("gvlma")) install.packages("gvlma")

library(caret)

library(caTools)
library(car)
library(quantmod)
library(MASS)
library(corrplot)

library(pder) # data
library(plm) # function: plm
library(stargazer)
library(Ecdat) # contains economic data
library(gmm)
library(dplyr)
library(magrittr)
library(lmtest)
library(tseries)
library(LINselect)
library(vars)
library(xtable)
library(StepReg)
library(corrplot)
library(tidycensus)
library(tidyverse)
library(plyr)
library(gvlma)
library(Metrics)
library(MLmetrics)
library(Hmisc)
```

Load data  
```{r}
setwd("/Users/shjour/Desktop")
m_data <- read.csv("ECO1400_MC.csv")

m_data$DATE <- as.Date(m_data$DATE)

#m_data <- m_data %>% mutate_at(c(25:25), as.numeric)
#m_data <- m_data %>% mutate_at(c(26:26), as.numeric)

# We drop the omitted data points.

m_data <- na.omit(m_data)

colnames(m_data)[1] <- "Date"

#m_data_lp <- m_data[-c(1,T10Y2YM,T10Y3MM)]
m_data_lp <- subset(m_data, 
                    select = -c(Date,T10Y2YM,T10Y3MM,GCCMXCHG))
m_data_lp <- log(m_data_lp)

m_data_log <- cbind(subset(m_data, 
                    select = c(Date,T10Y2YM,T10Y3MM,GCCMXCHG)), m_data_lp)

```

We first to check the correlation of our candiDate independent variables. To achieve this, we create the correlation matrix here.
```{r}

df_m_log <- m_data_log[-c(1)]

df_m_log.cor <- cor(df_m_log)
df_m_log.cor
#col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
#corrplot(df_m_log.cor, method="color", col=col(200),  
#         type="upper", order="hclust", 
#        addCoef.col = "black", # Add coefficient of correlation
#         tl.col="black", tl.srt=45, #Text label color and rotation
#         # Combine with significance
#         sig.level = 0.01, insig = "blank", 
#         # hide correlation coefficient on the principal diagonal
#         diag=FALSE 
#         )
corrplot(df_m_log.cor)
```

We drop the variable that has small correlation to the dependent variable MCOILWTICO, which is
T10Y2YM T10Y3MM M1SL MVIXHGH, M2SL, TB3MS, UMCSENT. Also, the rest of which have strong correlation for the independence. We drop IQ, MVIXOPN, MVIXLOW.

```{r}
m_data_log <- subset(m_data_log, 
                    select = -c(T10Y2YM,T10Y3MM,M1SL,MVIXHGH,IQ,MVIXOPN,MVIXLOW))
m_data_log
```

Now, we use the selected variables to do step-wise regression.
```{r}
m_data_mr_sw <- stepwise(MCOILWTICO ~ CCUSMA02RUM618N +CPIAUCSL +CPILFESL +CUSR0000SETA02 +INDPRO +IR +MCOILBRENTEU +MSACSR +POPTHM +UNRATE +MVIXADCLS +GCCMXCLS, data = m_data)

m_data_mr_sw

##
```

Thus, we choose MCOILBRENTEU, GCCMXCLS, MSACSR, CUSR0000SETA02, IR, UNRATE, CPILFESL, POPTHM ,CPIAUCSL.
```{r}
m_data_log <- subset(m_data_log, 
                    select = c(Date,MCOILWTICO, MCOILBRENTEU, GCCMXCLS, MSACSR, CUSR0000SETA02, IR, UNRATE, CPILFESL, POPTHM ,CPIAUCSL,MVIXADCLS))
```

Thus, we now try to use Multivariable Regression to predict the crude oil price.
```{r}
name_Y  <- "MCOILWTICO" # crude oil price
train_index <- c(1:round(0.7*nrow(m_data_log)) )
# train_index <- sample(1:nrow(m_data_log),round(0.7*nrow(m_data_log))) 
# row indeces of data in the training sample
train_data <- m_data_log[train_index,]
holdout_data <- m_data_log[-train_index,]
Y_holdout <- m_data_log[-train_index,name_Y]
X_holdout <- data.frame(m_data_log[-train_index,])
model_MR <- lm(MCOILWTICO ~ MCOILBRENTEU+ GCCMXCLS+ MSACSR+ CUSR0000SETA02+ IR+ UNRATE+ CPILFESL+ POPTHM+ CPIAUCSL+ MVIXADCLS, data=train_data)

vif_values <- vif(model_MR)           #create vector of VIF values
vif_values

# We drop the potential multilinearity variables.

model_MR <- lm(MCOILWTICO ~ MSACSR+ IR +UNRATE +CUSR0000SETA02+ CPIAUCSL+ MVIXADCLS, data=train_data)

vif_values <- vif(model_MR)           #create vector of VIF values
vif_values

summary(model_MR)

m_data_log <- subset(m_data_log, 
                    select = c(Date, MCOILWTICO, MSACSR, CUSR0000SETA02, IR, UNRATE ,CPIAUCSL,MVIXADCLS))

write.csv(m_data_log, "/Users/shjour/Desktop/select_m_data.csv", row.names = FALSE)
stargazer(model_MR, title="Results", no.space = TRUE, omit.stat=c("f", "ser"))

gvlma(model_MR)

```

We check the OLS assumptions for OLS model.
```{r}
par(mfrow=c(2,2))
plot(model_MR, which=1) #Residuals vs. Fitted, Linearity of the data
plot(model_MR, which=2) #the distribution of residuals across the model, Normality of residuals
plot(model_MR, which=3) #Homogeneity of variance
plot(model_MR, which=5)


```


Plot SR and predictions
```{r}
predict_MR <- predict(model_MR, newdata=holdout_data)
RMSE_MR <- mean((predict_MR - Y_holdout)^2)
MAE_MR <- mae(Y_holdout, predict(model_MR, newdata=holdout_data))
MAPE_MR <- MAPE(predict_MR, Y_holdout)
RMSE_MR
MAE_MR
MAPE_MR


predict_data <- cbind(X_holdout,predict_MR)
colors <- c("Training data"="darkgrey", "Holdout data"="green", "Regression"="blue")
ggplot() +
geom_point(data=train_data, aes(x=train_data[,"Date"], y=train_data[,name_Y], color="Training data")) +
geom_point(data=holdout_data, aes(x=holdout_data[,"Date"], y=holdout_data[,name_Y], color="Holdout data")) +
geom_point(data=predict_data, aes(x=predict_data[,"Date"], y=predict_MR, color="Regression")) +
geom_line(data=predict_data, aes(x=predict_data[,"Date"], y=predict_MR, color="Regression")) +
labs(x="Date", y="median value", title="") + scale_color_manual(name="", values=colors) +
theme_minimal() + theme(legend.position=c(0.12, 0.87), legend.title = element_blank())



```

